---
title: "Lab 06 - Functions and iterations"
author: "Kate Robinson"
date: "`r Sys.Date()`"
output: html_document
---


### Packages and Data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(ggplot2)
```

```{r diamon_rename}
diamonds <- diamonds %>% rename(
    depth_pct = depth,
    length_mm = x,
    width_mm = y,
    depth_mm = z,
    table_pct = table
  )
```


## Exercise 1

```{r}
# Do some preliminary investigations & cleaning

diamonds %>%
  count(color)

diamonds %>%
  count(cut)

diamonds %>%
  count(clarity)

diamonds %>%
  ggplot(aes(carat)) +
  geom_histogram(bins = 100)

diamonds <- diamonds %>%
  filter(carat <= 3)

diamonds %>%
  ggplot(aes(price)) +
  geom_histogram(bins = 100)

diamonds %>%
  ggplot(aes(length_mm)) +
  geom_histogram(bins = 100)

diamonds <- diamonds %>%
  filter(length_mm >= 2.5)

diamonds %>%
  ggplot(aes(width_mm)) +
  geom_histogram(bins = 100)

diamonds <- diamonds %>%
  filter(width_mm <= 20)

diamonds %>%
  ggplot(aes(depth_mm)) +
  geom_histogram(bins = 100)

diamonds <- diamonds %>%
  filter(depth_mm <= 10)

diamonds <- diamonds %>%
  filter(length_mm != 0 & depth_mm != 0 & width_mm != 0)

diamonds

```


## Exercise 2

```{r}

diamonds <- diamonds %>%
  mutate(table_mm = table_pct * width_mm / 100)

# Overall table_mm arithmetic mean?


diamonds %>%
  summarise(
    table_mm_mean = mean(table_mm)
  )



# table_mm arithmetic mean within each clarity category


diamonds %>%
  group_by(clarity) %>%
  summarise(
    table_mm_mean = mean(table_mm)
  )



```

_As clarity improves, the table percentage mean decreases._

```{r}
# calculate the arithmetic mean across all numeric variables

diamonds %>%
  summarise(
    avg_carat = mean(carat),
    avg_depth_pct = mean(depth_pct),
    avg_table_pct = mean(table_pct),
    avg_price = mean(price),
    avg_length_mm = mean(length_mm),
    avg_width_mm = mean(width_mm),
    avg_depth_mm = mean(depth_mm),
    avg_table_mm = mean(table_mm)
  )



```

## Exercise 3

```{r}
# Write gmean() command

gmean <- function(x){
  logsum <- 0
  for(i in x){
    logsum <- logsum + log(i)
  }
  exp(1/length(x)*logsum)
}

```

## Exercise 4

```{r, error = TRUE}

test_data <- list(
  test1 = c(2.1, 3.8, 4.2),
  test2 = c(1, 10, 100, 1000),
  test3 = c(0, 1, 4),
  test4 = c(0.38,  0.94, -1.56),
  test5 = c(TRUE, TRUE),
  test6 = c("6", "7", "8")
)


# Create for loop to test gmean() on the above examples
for(item in test_data){
  print(gmean(item))
}

# doesn't work for test5 as non-numeric values, doesn't work for test6 as character values so all taken as 1

```


## Exercise 5

```{r}
# Copy gmean() from Ex 3 and edit it to check the input


gmean <- function(x){
  if(any(x <= 0)){
      warning("Geometric mean for positive values only")
      return(NaN)
  } else if(any(!is.numeric(x))){
      warning("Geometric mean for numeric values only")
      return(NaN)
    }

  logsum <- 0
  for(i in x){
    logsum <- logsum + log(i)
  }
  exp(1/length(x)*logsum)
}



# Copy your code from Ex 4 to test your new command


for(item in test_data){
  print(gmean(item))
}



```

## Exercise 6

```{r}
# replicate the for loop from Ex 4 using the map (or a similar) function
map_dfr(test_data, gmean) # could also use map or do with rows instead of columns, etc.

```


## Exercise 7

```{r}
# Overall table_mm arithmetic mean, median, and geometric mean?

diamonds %>%
  summarise(
    table_mm_mean = mean(table_mm),
    table_mm_median = median(table_mm),
    table_mm_geommean = gmean(table_mm)
  )




# Arithmetic mean, median, and geometric mean of table_mm within each clarity category

diamonds %>%
  group_by(clarity) %>%
  summarise(
    table_mm_mean = mean(table_mm),
    table_mm_median = median(table_mm),
    table_mm_geommean = gmean(table_mm)
  )




```

_As before we see the table_mm decreases with increasing clarity - this is the case with all types of mean._


## Exercise 8 (optional)

```{r}
# Create gmean2() that computes the geometric mean 
#   that uses the original definition
library(pracma)

gmean2 <- function(x){
  if(any(x <= 0)){
      warning("Geometric mean for positive values only")
      return(NaN)
  } else if(any(!is.numeric(x))){
      warning("Geometric mean for numeric values only")
      return(NaN)
    }
  mult <- 1
  for(val in x){
    mult <- mult * val
  }
  nthroot(mult, length(x))
}

map_dfr(test_data, gmean2)
#fine here

diamonds %>%
  summarise(
    table_mm_geommean = gmean2(table_mm)
  )
#so many values it doesn't work

```


## Exercise 9 (optional)

```{r}
# Create a function that computes the harmonic mean


hmean <- function(x){
  if(any(x <= 0)){
      warning("Harmonic mean for positive values only")
      return(NaN)
  } else if(any(!is.numeric(x))){
      warning("Harmonic mean for numeric values only")
      return(NaN)
  }
  recipx <- c()
  for(i in x){
    recipx <- c(recipx, 1/i)
  }
  recipmean <- mean(recipx)
  1/recipmean
}

map_dfr(test_data, hmean)

diamonds %>%
  summarise(
    table_mm_harmmean = hmean(table_mm)
  )


```

## Exercise 10 (optional)

_Do derivation on paper, unless you know how to write LaTeX code_
